### 2019-01-15 submission/manuscript.tex
Fresh make write.paper

\documentclass[11pt,]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1.0in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

%%% Change title format to be more compact
\usepackage{titling}

% Create subtitle command for use in maketitle
\newcommand{\subtitle}[1]{
  \posttitle{
    \begin{center}\large#1\end{center}
    }
}

\setlength{\droptitle}{-2em}

  \title{}
    \pretitle{\vspace{\droptitle}}
  \posttitle{}
    \author{}
    \preauthor{}\postauthor{}
    \date{}
    \predate{}\postdate{}

\usepackage{helvet} % Helvetica font
\renewcommand*\familydefault{\sfdefault} % Use the sans serif version of the font
\usepackage[T1]{fontenc}

\usepackage[none]{hyphenat}

\usepackage{setspace}
\doublespacing
\setlength{\parskip}{1em}

\usepackage{lineno}

\usepackage{pdfpages}

\usepackage{amsmath}

\usepackage{mathtools}

\begin{document}

\vspace{25mm}
\begin{center}

\textbf{\huge Assessing the differences in 16S rRNA gene sequence data that are due to choice of DNA polymerase and number of rounds of PCR}

\vspace{50mm}

\textbf{Running title:} Quantifying the choice of DNA polymerase

\vspace{25mm}

Marc A Sze${^1}$ and Patrick D Schloss${^1}$${^\dagger}$

\vspace{20mm}

$\dagger$ To whom correspondence should be addressed: pschloss@umich.edu

$1$ Department of Microbiology and Immunology, University of Michigan, Ann Arbor, MI

\end{center}

\newpage
\linenumbers

\hypertarget{abstract}{%
\subsection{Abstract}\label{abstract}}

PCR amplification of 16S rRNA genes is a critical, yet under appreciated
step in the generation of sequence data to describe the taxonomic
composition of microbial communities. Numerous factors in the design of
PCR can impact the sequencing error rate, the abundance of chimeric
sequences, and the degree to which the fragments in the product
represent their abundance in the original sample. The choice of PCR
cycling conditions are known to impact the outcome of PCR and downstream
inferences. We compared the performance of high fidelity polymerases and
varying number of rounds of amplification when amplifying a mock
community with known sequences and community structure and human stool
samples. Although it is impossible to derive specific recommendations,
we did observe general trends. Namely, using a polymerase with the
highest possible fidelity and minimizing the number of rounds of PCR
reduced the sequencing error rate, fraction of chimeric sequences, and
bias. Evidence of bias at the sequence level was subtle and could not be
ascribed to the fragments' fraction of bases that were guanines or
cytosines. When analyzing the mock community data, the amount that the
community deviated from the expected composition increased with rounds
of PCR. The effect was inconsistent for human stool samples. Overall the
results underscore the difficulty of comparing sequence data that are
generated by different PCR protocols; however, the results indicate that
the variation in human stool samples is generally larger than that
introduced by the choice of polymerase or number of rounds of PCR.

\hypertarget{importance}{%
\subsubsection{Importance}\label{importance}}

The growth in interest in characterizing microbial communities from
diverse communities has exploded with the steep decline in sequencing
costs. A significant amount of effort has gone into understanding the
error profiles of DNA sequencers; however, little work has been done to
understand the effects of choices that researchers make to generate the
PCR amplicons that are sequenced. Here we quantified the effects of the
choice of polymerase and number of PCR cycles on the quality of
downstream data. Overall, we found that these choices can have a
profound impact on the way that a microbial community is represented in
the sequence data. Although the effects are relatively small compared to
the variation in human stool samples, care should be taken to use
polymerases with the highest possible fidelity and to minimize the
number of rounds of PCR.

\newpage

\hypertarget{introduction}{%
\subsection{Introduction}\label{introduction}}

16S rRNA gene sequencing is a powerful and widely used tool for
surveying the structure of microbial communities (1--3). This approach
has exploded in popularity with the advent of high throughput sequencing
where it is possible to characterize numerous samples with thousands of
sequences per sample. Numerous factors can impact how a natural
community is represented by the sequencing data including the method of
acquiring samples (4--8), storage conditions (4--6, 9--12), extraction
methods (13), amplification conditions (8, 14, 15), sequencing method
(15--17), and analytical pipeline (15, 18--20). The increased sampling
depth that is now available relative to previous Sanger-based methods is
expected to compound the impacts of an investigator's choices and the
interpretation of their results.

One step in the generation of 16S rRNA gene sequence data that has been
long known to have a significant impact on the description of microbial
communities is the choice of conditions for PCR amplification (8, 14,
15). Factors such as the choice of primers have an obvious impact on
which populations will be amplified (18, 21). However, a variety of PCR
artifacts can also impact the perception of a community including the
formation of chimeras (14, 22--24), misincorporation of nucleotides (23,
25, 26), preferential amplification of populations leading to bias (24,
27--33), and accumulation of random amplification events leading to PCR
drift (24, 27, 32, 34). Many bioinformatic tools have been developed to
identify chimeras, there are significant sensitivity and specificity
tradeoffs (14, 35). Others have attempted to account for PCR bias using
modeling approaches (29, 36); however, these have been developed for
idealized situations. Laboratory-based solutions to minimize chimera
formation have also been proposed such as minimizing the amount of
template DNA in the PCR, minimizing the number of rounds of PCR,
minimizing the amount of shearing in the template DNA, and using DNA
polymerases that have a proof-reading ability (14, 23). To minimize PCR
drift, some investigators will pool technical replicate PCRs hoping to
average out the drift (34). Other factors that have been shown to impact
the formation of PCR artifacts are outside the control of an
investigator including the fraction of DNA bases that are guanines or
cytosines, the variation in the length of the targeted region across the
community, the sequence of the DNA that flanks the template, and the
genetic diversity of the community (28, 30--33). Early investigations of
the factors that lead to the formation of PCR artifacts focused on
analyzing binary mixtures of genomic DNA and 16S rRNA gene fragments to
explore PCR biases and chimera formation. Although these studies were
instrumental in forcing researchers to be cautious about the
interpretation of their results, we have a poor understanding of how
these factors affect the formation of PCR artifacts in more complex
communities.

The influence that the choice of DNA polymerase has on the formation of
PCR artifacts has not been well studied. There has been recent interest
in how the choice of the hypervariable region and data analysis
pipelines impact the sequencing error rate (15, 18--20); however, these
studies use the same DNA polymerase in the PCR step and implicitly
assume that the rate of nucleotide misincorporation from PCR are
significantly smaller than those from the sequencing phase. There has
been more limited interest in the impact that DNA polymerase choice has
on the formation of chimeras (23, 37). A recent study found differences
in the number of OTUs and chimeras between normal and high fidelity DNA
polymerases (37). The authors of this study could reduce the difference
between two polymerases by optimizing the annealing and extension steps
within the PCR protocol (37). Yet this optimization was specific for the
community they were analyzing (i.e.~captive and semi-captive red-shanked
doucs) and assumed that if the two polymerases generate the same
community structure that the community structure was correct. In fact,
the community structure generated by both methods was not free of
artifacts, but has the same artifacts. A challenge in these types of
experiments is having a priori knowledge of the true community
representation. A mock community with known composition allows
researchers to quantify the sequencing error rate, fraction of chimeras,
and bias (19); however, mock communities have a limited phylogenetic
diversity relative to natural communities. Natural communities, in
contrast, have an unknown community composition making absolute
measurements impossible. They can be used to validate results from mock
communities and to understand the relative impacts of artifacts on the
ability to differentiate biological and methodological sources of
variation. Given the large number of DNA polymerases available to
researchers it is unlikely that a specific recommendation is possible.
Rather, the development of general best practices and understanding the
impact of PCR artifacts on an analysis are desired.

This study investigated the impact of choice of high fidelity DNA
polymerase and the number of rounds of amplification on the formation of
PCR artifacts using a mock community and human stool samples. These PCR
artifacts included (i) the effect of the polymerase on the error rate of
the bases represented in the final sequences, (ii) the fraction of
sequences that appeared to be chimeras and the ability to detect those
chimeras, (iii) the bias of preferentially amplifying one fragment over
another in a mixed pool of templates, and (iv) inter-sample variation in
community structure of samples amplified with the same polymerase across
the amplification process. To characterize these factors we sequenced a
mock community of 8 organisms with known sequences and community
structure and human fecal samples with unknown sequences and community
structures. We sequenced the V4 region of the 16S rRNA genes from a mock
community by generating paired 250 nt reads on the Illumina MiSeq
platform. This region and sequencing approach was used because it has
been shown to result in a relatively low sequencing error rate and is a
widely used protocol. To better understand the impact of DNA polymerase
choice on PCR artifacts, we selected five high fidelity DNA polymerases
and amplified the communities using 15, 20, 25, 30, and 35 rounds of
amplification (Table 1). Collectively, our results suggest that the
number of rounds and to a lesser degree the choice of DNA polymerase
used in PCR impact the sequence data, the effects are consistent and are
smaller than the biological differences between individuals.

\newpage

\hypertarget{results}{%
\subsection{Results}\label{results}}

\textbf{\emph{Sequencing errors vary by the number of cycles and the DNA
polymerase used in PCR.}} The presence of sequence errors can confound
the ability to accurately classify 16S rRNA gene sequences and group
sequences into operational taxonomic units (OTUs). More importantly,
sequencing errors themselves can alter the representation of the
community. Therefore, it is important to minimize the number of
sequencing errors. Using a widely-used approach that generates the
lowest reported error rate, we quantified the error rate by sequencing
the V4 region of the 16S rRNA genes from an 8 member mock community. We
also removed any contigs that were at least three bases more similar to
a chimera of two references than to a single reference sequence (18, 19,
38). Regardless of the polymerase, the error rate increased with the
number of rounds of amplification (Figure 1). Using 30 rounds of PCR is
a common approach across diverse types of samples. Among the data
generated using 30 rounds of PCR the Accuprime polymerase had the
highest error rate (i.e.~0.124\%) followed by the Platinum
(i.e.~0.094\%), Phusion (i.e.~0.064\%), Kappa (i.e.~0.062\%), and Q5
(i.e.~0.060\%) polymerases (Figure 1). When we applied the
pre.clustering denoising step, which merges the counts of reads within 2
nt of a more abundant sequence, the error rates dropped considerably
such that the Platinum polymerase had the highest error rate
(i.e.~0.014\%) followed by the Accuprime (i.e.~0.012\%), Q5
(i.e.~0.0053\%), Phusion (i.e.~0.0049\%), and Kappa (i.e.~0.0049\%)
polymerases (Figure 1). Although specific recommendations are difficult
to make because the impact of the actual community structure and
concentration of the initial DNA template are likely to have an impact
on the results, it is clear that using as few PCR cycles as necessary
and a polymerase with the lowest possible error rate is a good guide to
minimizing the impact of polymerase on the error rate.

\textbf{\emph{The fraction of sequences identified as being chimeric
varies by the number of cycles and the DNA polymerase used in PCR.}}
Chimeric PCR products can significantly confound downstream analyses.
Although numerous bioinformatic tools exist to identify and remove
chimeric sequences with high specificity, their sensitivity is
relatively low and can be reduced by the presence of sequencing errors.
We identified those contigs that were at least three bases more similar
to a chimera of two references than to a single reference sequence. As
expected from previous studies, the number of chimeras increased with
rounds of amplification (Figure 2A). Interestingly, the fraction of
chimeric sequences from the mock community varied by the type of
polymerase used. After 30 rounds of PCR, the Platinum polymerase had the
highest chimera rate (i.e.~18.2\%) followed by the Q5 (i.e.~8.1\%),
Phusion (i.e.~7.5\%), Kappa (i.e.~2.3\%), and Accuprime (i.e.~0.9\%)
polymerases. We used the UChime algorithm to detect chimeras in our
pre-clustered mock community data and calculate the algorithm's
sensitivity and specificity (Figure 2A). For all polymerases, the
specificity was above 95.7\% and showed a weak association with the
number of cycles used. There was considerable inter-polymerase and
inter-round of amplification variation in the sensitivity of UChime to
detect the chimeras from the mock community. This suggested that the
residual error rate after pre-clustering the sequence data did not
compromise the sensitivity. The sensitivity of UChime varied between 50
and 87.0\%. The generalizability of these results is limited because we
used a single mock community with limited genetic diversity. Although we
did not know the true chimera rate for our four human stool samples, we
were able to calculate the fraction of sequences that UChime identified
as being chimeric (Figure 2B). These results followed those from the
mock communities; additional rounds of amplification significantly
increased the rate of chimeras and there was variation between the
polymerases that we used. Although it was not possible to identify the
features of a polymerase that resulted in higher rates of chimeras, it
is clear that using the smallest number of PCR cycles possible will
minimize the impacts of chimeras.

\textbf{\emph{At the sequence level, PCR amplification bias is subtle.}}
Since researchers began using PCR to sequence 16S rRNA genes there has
been concern that amplifying fragments from a mixed template pool could
lead to a biased representation in the pool of products and would
confound downstream analyses. The mock community was generated by mixing
equal amounts of genomic DNA from each of the 8 bacteria resulting in
uneven representation of the \emph{rrn} operons across the bacteria. The
vendor of the mock community subjects each lot of genomic DNA to shotgun
sequencing to more accurately quantify the actual abundance of each
organism in the community. We compared the expected relative abundance
of the 16S rRNA genes from each bacterium in the mock community to the
data we generated across rounds of amplification and polymerase (Figure
3). Interestingly, for some bacteria, their representation became less
biased with additional rounds of PCR (e.g.~\emph{L. fermentum}), while
others became more biased (e.g.~\emph{E. faecalis}), and others had
little change (e.g.~\emph{B. subtilis}). The percentage of bases in the
V4 region that were guanines or cytosines was not predictive of the
amount of bias. Across the strains there was no variation in the length
of their V4 regions and they each had the same sequence in the region
that the primers annealed. One of the bacteria represented in the mock
community, \emph{S. enterica}, had 6 identical copies of the V4 region
and 1 copy that differed from those by one nucleotide. The dominant copy
had a thymidine and the rare copy had a guanine. We used the sequence
data to calculate the ratio of the dominant to rare variants from
\emph{S. enterica} (Figure S1). The Accuprime, Phusion, Platinum, and Q5
polymerases converged to a ratio of 5.4; however, the ratio for the
Kappa polymerase varied between 6.1 and 6.8 for 25 to 35 rounds of PCR.
Given the subtle nature of the variation in the relative abundances of
each 16S rRNA gene fragment, it was not possible to create generalizable
rules that would explain the bias and the previous observation regarding
the guanine and cytosine relative abundance was not observed with our
data.

\textbf{\emph{At the community level, the effects of PCR amplification
bias grow with additional rounds of PCR.}} Because the distance between
samples and across rounds of PCR could be artificially inflated due to
sequencing errors and chimeras, we used analyzed the alpha and beta
diversity of the mock community data at different phases of the sequence
curation pipeline (Figure 4). First, we removed the chimeras from the
mock community data as described above and mapped the individual reads
to the OTUs that the 16S rRNA gene fragments would cluster into if there
were no sequencing errors. This gave us a community distribution that
reflected the distribution following PCR without any artifacts. Alpha
diversity metrics such as richness and Shannon and Inverse Simpson
diversity metrics increased with the number of rounds of PCR for all
polymerases except for the Kappa polymerase, where the richness remained
constant but the diversity decreased (Figure 4A). These data suggest
that PCR had the effect of making the community distribution more even
than it was originally. Next, we used the true sequence diversity, but
removed chimeras based on mapping reads to all possible chimeras between
reference sequences, and clustered the reads to OTUs. The same trends
were observed as with perfect sequence data except that the richness and
diversity metrics trended higher (Figure 4A). Finally, we used the
observed sequence data and used the UCHIME algorithm to identify
chimeras. Again, we observed the same trends as with the other analyses
except the richness and diversity metrics trended higher (Figure 4A).
These comparisons demonstrated that although the bias at the sequence
level was subtle, PCR does have a bias that is exacerbated by errors and
chimeras. When we measured the Bray-Curtis distance between the
communities observed after 25 rounds of amplification and those at 30
and 35 rounds of amplification the distances between 25 and 35 rounds
were higher than between 25 and 30 rounds for each of the polymerases by
an average of 0.022 units (Figure 4B). The Platinum polymerase varied
the most across rounds of amplification (25 vs 30 rounds: 0.12; 25 vs 35
rounds: 0.15). Although the distances between samples were small, the
ordination of these distances showed a clear change in community
structure with increasing rounds of PCR (Figure 4C). This observation
was supported by our statistical analysis, which revealed that the
number of rounds of PCR (R\textsuperscript{2}=0.21, P\textless{}0.001)
was a larger factor than the choice of polymerase
(R\textsuperscript{2}=0.20, P\textless{}0.001). These results
demonstrate that subtle differences in relative abundances can have an
impact on overall community structure. This variation underscores the
importance of only comparing sequence data that have been generated
using the same PCR conditions.

\textbf{\emph{The choice of polymerase or the number of rounds of
amplification have little impact on the relative interpretation of
community-wide metrics of diversity.}} The biases that we observed at
the population and community levels using mock community data appeared
to be small relative to the expected differences between biological
samples. To study this further, we calculated alpha and beta-diversity
metrics using the human stool samples for each of the polymerases and
rounds of amplification. We calculated the number of observed OTUs,
Shannon diversity, inverse Simpson diversity index for each condition
and donor (Figure 5A). Although there were clear differences between
conditions, the relative ordering of the stool samples did not
meaningfully vary across conditions. When we characterized the variation
between rounds of amplification using human stool samples, the distance
between the 25 and 30 rounds and 25 and 35 rounds varied considerably
between samples and polymerases (Figure 5B). In general the inter-round
variation was lowest for the data generated using the Kappa and
Accuprime polymerases. The data generated using the Platinum polymerase
was consistent across rounds, but it was more biased than these
polymerases. Considering the average distance across the four samples
varied between 0.39 and 0.56, regardless of the polymerases and number
of rounds of amplification, any bias due to amplification is unlikely to
obscure community-wide differences between samples. In support of this
was our principle coordinates analysis of the Bray-Curtis distances,
which revealed distinct clusters by donor (Figure 5C). Within each
cluster there were no obvious patterns related to the polymerase or
number of rounds of PCR. However, our statistical analysis using adonis
revealed statistically significant differences in the community
structures with the subject explaining the most variation
(R\textsuperscript{2}=0.79, P\textless{}0.001), followed by the number
of rounds of PCR (R\textsuperscript{2}=0.044, P\textless{}0.001) and the
choice of polymerase (R\textsuperscript{2}=0.033, P\textless{}0.001).
Together, these results indicate that for a coarse analysis of
communities, the choice of number of rounds of amplification or
polymerase are not important, but that they must be consistent across
samples. It is difficult to develop a specific recommendation based on
the level of bias across rounds of PCR or polymerases; however, the
general suggestion is to use as few rounds of amplification as possible.

\textbf{\emph{There is little evidence of a relationship between
polymerase or number of rounds of amplification on PCR drift.}} There
have been concerns that the same template DNA subjected to the same PCR
conditions could result in different representations of communities
because of random drift. To test this, we determined the average
Bray-Curtis distance between replicate reactions using the same
polymerase and number of rounds of amplification (Figure 6). Using the
mock community data there were no obvious trends. The average
Bray-Curtis distance within a set of conditions varied by 0.063 to 0.11
units. Although we did not obtain replicates of each of the stool
samples, the intra-sample variation for each set of conditions was
consistent and varied between 0.50 and 0.56 units. These data suggest
that amplicon sequencing is robust to random variation in amplification
and that any differences are likely to be smaller than what is
considered biologically relevant.

\newpage

\hypertarget{discussion}{%
\subsection{Discussion}\label{discussion}}

Our results suggest that the number of rounds of PCR and to a lesser
degree the choice of DNA polymerase impact the analysis of 16S rRNA gene
sequence data from bacterial communities. Although it was not possible
to make direct connections between PCR conditions and specific sources
of bias, we were able to identify general recommendations that reduce
the amount of error, chimera formation, and bias. Researchers should
strive to minimize the number of rounds of PCR and should use a high
fidelity polymerase. Although specific PCR conditions impact the precise
interpretation of the data, the effects were consistent and were smaller
than the biological differences between individuals. Based on these
observations, within a study, amplicons must be generated by consistent
protocols to yield meaningful comparisons. When comparing across
studies, values like richness, diversity, and relative abundances must
be made in relative and not absolute terms.

The observed sequencing error rates and alpha diversity metrics followed
the manufacturers' measurements of their polymerases' fidelity (Figure
1). Accuprime and Platinum have fidelity that are approximately 10-times
higher than that of Taq whereas the fidelity of Phusion, Q5, and Kappa
are more than 100 times higher. Among these polymerases, the Kappa
polymerase resulted in the the lowest error rate, lowest chimera rate,
and least bias across rounds of PCR. Considering polymerase development
is an active area of commercial development and it likely that new
polymerases will come to market, it is important for researchers to
understand how changing the polymerase impacts downstream analyses for
their type of samples.

Over the past 20 years, a large literature has attempted to document
various PCR biases and underscores the fact that data based on
amplification of DNA from a mixed community are not a true
representation of the actual community. In addition to obvious biases
imposed by primer selection, other factors inherent in PCR can influence
the representation of communities. Factors that can lead to preferential
amplification of one fragment over another have included guanine and
cytosine composition, length, flanking DNA composition, amount of DNA
shearing, and number of rounds of PCR (24, 27--33). In addition,
environmental and reagent contaminants can also have a significant
impact on the analysis of low biomass samples (39). Less well understood
is the effect of phylogenetic diversity on bias and chimera formation.
Communities with low phylogenetic diversity may be more prone to chimera
formation since chimeras are more likely to form among closely related
sequences (14, 35). The interaction of these various influences on PCR
artifacts are complex and difficult to tease apart. Minimizing the level
of DNA shearing and using the fewest number of rounds of PCR with a
polymerase that has the highest possible fidelity are strategies that
can be employed to minimize the formation of chimeras. Although care
should always be taken when choosing a polymerase for 16S rRNA gene
sequencing, our observations show that the differences between a variety
of polymerases are smaller than the actual biological variation in fecal
communities between individuals. Therefore, if the biological signal of
interest is similar to differences in fecal bacterial communities found
between individuals, then the type of high fidelity DNA polymerase used
will only minimally change the results.

Even with these strategies it is impossible to remove all PCR artifacts.
Beyond the imperfections of the best polymerases, sometimes difficult to
lyse organisms require stringent lysis steps and low biomass samples
require additional rounds of PCR. A host of bioinformatics tools are
available for removing residual sequencing errors (18, 40--42). These
tools struggle to correctly differentiate true biological diversity
(i.e.~Amplicon Sequencing Variants or ASVs) and PCR or sequencing
errors. Although these methods remove residual errors, they also risk
splitting intragenomic variants into separate ASVs, merging 16S rRNA
gene sequences from different taxa into the same ASV. Other tools are
available for removing chimeras (14, 35) where there is a trade off
between the sensitivity of detecting chimeras and the specificity of
correctly calling a sequence a chimera. In recent years, parameters for
these algorithms have been changed to increase their sensitivity with
little evaluation of the effects on the specificity of the algorithms
(40, 42). Others recommend removing any read that has an abundance below
a specified threshold (e.g.~removing all sequences that only appear
once) (20, 40--42). This method must be approached with caution as such
approaches are likely to introduce a different bias of the community
representation and ignore the fact that artifacts may be quite abundant.
Ultimately, researchers must test their hypotheses with multiple methods
to validate the claims they reach with any one method (43). All methods
have biases and limitations and we must use complementary methods to
obtain robust results.

\newpage

\hypertarget{materials-methods}{%
\subsection{Materials \& Methods}\label{materials-methods}}

\textbf{\emph{Mock community.}} The ZymoBIOMICS\textsuperscript{TM}
Microbial Community DNA Standard (Zymo, CA, USA) was used for mock
communities and the bacterial component was made up of \emph{Pseudomonas
aeruginosa}, \emph{Escherichia coli}, \emph{Salmonella enterica},
\emph{Lactobacillus fermentum}, \emph{Enterococcus faecalis},
\emph{Staphylococcus aureus}, \emph{Listeria monocytogenes}, and
\emph{Bacillus subtilis} at equal genomic DNA abundance
(\url{https://web.archive.org/web/20171217151108/http://www.zymoresearch.com:80/microbiomics/microbial-standards/zymobiomics-microbial-community-standards}).
The actual relative abundance for each bacterium was obtained from
Zymo's certificate of analysis for the lot (Lot: ZRC187325), which they
determined using shotgun metagenomic sequencing
(\url{https://github.com/SchlossLab/Sze_PCRSeqEffects_mSphere_2019/data/references/ZRC187325.pdf}).

\textbf{\emph{Human samples.}} Fecal samples were obtained from 4
individuals who were part of an earlier study (44). These samples were
collected using a protocol approved by the University of Michigan
Institutional Review Board. For this study, the samples were
de-identified. DNA was extracted from the fecal samples using the
MOBIO\textsuperscript{TM} PowerMag Microbiome RNA/DNA extraction kit
(now Qiagen, MD, USA).

\textbf{\emph{PCR protocol.}} Five high fidelity DNA polymerases were
tested including AccuPrime\textsuperscript{TM} (ThermoFisher, MA, USA),
KAPA HIFI (Roche, IN, USA), Phusion (New England Biolabs, MA, USA),
Platinum (ThermoFisher, MA, USA), and Q5 (New England Biolabs, MA, USA).
Manufacturer recommendations were followed except for the annealing and
extension times, which were selected based on previously published
protocols (18, 37). Primers targeting the V4 region of the 16S rRNA gene
were used with modifications to generate MiSeq amplicon libraries (18)
(\url{https://github.com/SchlossLab/MiSeq_WetLab_SOP/}). The number of
rounds of PCR used for each sample and polymerase 15 and increased by 5
rounds up to 35 cycles. Insufficient PCR product was generated using 15
rounds and has not been included in our analysis.

\textbf{\emph{Library generation and sequencing.}} Each PCR condition
(i.e.~combination of polymerase and number of rounds of PCR) were
replicated four times for the mock community and one time for each fecal
sample. Libraries were generated as described previously (18)
(\url{https://github.com/SchlossLab/MiSeq_WetLab_SOP/}). The libraries
were sequenced using the Illumina MiSeq sequencing platform to generate
paired 250-nt reads.

\textbf{\emph{Sequence processing.}} The mothur software program (v
1.41) was used for all sequence processing steps (45). The protocol has
been previously published (18)
(\url{https://www.mothur.org/wiki/MiSeq_SOP}). Briefly, paired reads
were assembled using mothur's make.contigs command to correct errors
introduced by sequencing (18). Any assembled contigs that contained an
ambiguous base call, mapped to the incorrect region of the 16S rRNA
gene, or appeared to be a contaminant were removed from subsequent
analyses. Sequences were further denoised using mothur's pre.cluster
command to merge the counts of sequences that were within 2 nt of a more
abundant sequence. The VSEARCH implementation of UCHIME was used to
screen for chimeras (35, 46). At various stages in the sequence
processing pipeline for the mock community data, the mothur seq.error
command was used to quantify the sequencing error rate as well as the
true chimera rate. This command uses the true sequences from the mock
community to generate all possible chimeras and removes any contigs that
were at least three bases more similar to a chimera than to a reference
sequence. The command then counts the number of substitutions,
insertions, and deletions in the contig relative to the reference
sequence and reports the error rate without the inclusion of chimeric
sequences (19). The reference sequences and operon copy number for each
bacterium were obtained from the ZymoBIOMICS\textsuperscript{TM}
Microbial Community DNA Standard protocol
(\url{https://web.archive.org/web/20181221151905/https://www.zymoresearch.com/media/amasty/amfile/attach/_D6305_D6306_ZymoBIOMICS_Microbial_Community_DNA_Standard_v1.1.3.pdf}).
Sequences were assigned to operational taxonomic units (OTUs) at a
threshold of 3\% dissimilarity using the OptiClust algorithm (47). To
adjust for unequal sequencing when measuring alpha and beta diversity,
all samples were rarefied to 1,000 sequences 1,000 times for downstream
analysis. The Good's coverage for the samples at this level of sampling
was routinely greater than 95\%.

\textbf{\emph{Statistical analysis.}} All analysis was done with the R
(v 3.5.0) software package (48). Data transformation and graphing were
completed using the tidyverse package (v 1.2.1) (49). The distance
matrix data was analyzed using the adonis function within the vegan
package (v 2.5.3) (50).

\textbf{\emph{Reproducible methods.}} The data analysis code for this
study can be found at
\url{https://github.com/SchlossLab/Sze_PCRSeqEffects_mSphere_2019}. The
raw sequences are available at the SRA (Accession SRP132931).

\newpage

\hypertarget{acknowledgements}{%
\subsection{Acknowledgements}\label{acknowledgements}}

The authors thank the study participants in ERIN whose samples were
utilized. We also would like to thank Judy Opp and April Cockburn for
their effort in sequencing the samples as part of the Microbiome Core
Facility at the University of Michigan. Additional thanks to members of
the Schloss lab and Dr.~Marcy Balunas for reading earlier drafts of the
manuscript and providing helpful critiques. Salary support for Marc A.
Sze came from the Canadian Institute of Health Research and NIH grant
UL1TR002240. Salary support for Patrick D. Schloss came from NIH grants
P30DK034933 and 1R01CA215574.

\newpage

\hypertarget{references}{%
\subsection{References}\label{references}}

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-Gilbert2018}{}%
1. \textbf{Gilbert JA}, \textbf{Jansson JK}, \textbf{Knight R}. 2018.
Earth microbiome project and global systems biology. mSystems
\textbf{3}.
doi:\href{https://doi.org/10.1128/msystems.00217-17}{10.1128/msystems.00217-17}.

\leavevmode\hypertarget{ref-HMP2012}{}%
2. \textbf{Consortium HM}. 2012. Structure, function and diversity of
the healthy human microbiome. Nature \textbf{486}:207--214.
doi:\href{https://doi.org/10.1038/nature11234}{10.1038/nature11234}.

\leavevmode\hypertarget{ref-Schloss2016}{}%
3. \textbf{Schloss PD}, \textbf{Girard RA}, \textbf{Martin T},
\textbf{Edwards J}, \textbf{Thrash JC}. 2016. Status of the archaeal and
bacterial census: An update. mBio \textbf{7}.
doi:\href{https://doi.org/10.1128/mbio.00201-16}{10.1128/mbio.00201-16}.

\leavevmode\hypertarget{ref-Luo2016}{}%
4. \textbf{Luo T}, \textbf{Srinivasan U}, \textbf{Ramadugu K},
\textbf{Shedden KA}, \textbf{Neiswanger K}, \textbf{Trumble E},
\textbf{Li JJ}, \textbf{McNeil DW}, \textbf{Crout RJ}, \textbf{Weyant
RJ}, \textbf{Marazita ML}, \textbf{Foxman B}. 2016. Effects of specimen
collection methodologies and storage conditions on the short-term
stability of oral microbiome taxonomy. Applied and Environmental
Microbiology \textbf{82}:5519--5529.
doi:\href{https://doi.org/10.1128/aem.01132-16}{10.1128/aem.01132-16}.

\leavevmode\hypertarget{ref-Bassis2017}{}%
5. \textbf{Bassis CM}, \textbf{Nicholas M. Moore}, \textbf{Lolans K},
\textbf{Seekatz AM}, \textbf{Weinstein RA}, \textbf{Young VB},
\textbf{Hayden MK}. 2017. Comparison of stool versus rectal swab samples
and storage conditions on bacterial community profiles. BMC Microbiology
\textbf{17}.
doi:\href{https://doi.org/10.1186/s12866-017-0983-9}{10.1186/s12866-017-0983-9}.

\leavevmode\hypertarget{ref-Gorzelak2015}{}%
6. \textbf{Gorzelak MA}, \textbf{Gill SK}, \textbf{Tasnim N},
\textbf{Ahmadi-Vand Z}, \textbf{Jay M}, \textbf{Gibson DL}. 2015.
Methods for improving human gut microbiome data by reducing variability
through sample processing and storage of stool. PLOS ONE
\textbf{10}:e0134802.
doi:\href{https://doi.org/10.1371/journal.pone.0134802}{10.1371/journal.pone.0134802}.

\leavevmode\hypertarget{ref-Dominianni2014}{}%
7. \textbf{Dominianni C}, \textbf{Wu J}, \textbf{Hayes RB}, \textbf{Ahn
J}. 2014. Comparison of methods for fecal microbiome biospecimen
collection. BMC Microbiology \textbf{14}:103.
doi:\href{https://doi.org/10.1186/1471-2180-14-103}{10.1186/1471-2180-14-103}.

\leavevmode\hypertarget{ref-BautistadelosSantos2016}{}%
8. \textbf{Santos QMB-d los}, \textbf{Schroeder JL}, \textbf{Blakemore
O}, \textbf{Moses J}, \textbf{Haffey M}, \textbf{Sloan W}, \textbf{Pinto
AJ}. 2016. The impact of sampling, PCR, and sequencing replication on
discerning changes in drinking water bacterial community over diurnal
time-scales. Water Research \textbf{90}:216--224.
doi:\href{https://doi.org/10.1016/j.watres.2015.12.010}{10.1016/j.watres.2015.12.010}.

\leavevmode\hypertarget{ref-Sinha2015}{}%
9. \textbf{Sinha R}, \textbf{Chen J}, \textbf{Amir A}, \textbf{Vogtmann
E}, \textbf{Shi J}, \textbf{Inman KS}, \textbf{Flores R},
\textbf{Sampson J}, \textbf{Knight R}, \textbf{Chia N}. 2015. Collecting
fecal samples for microbiome analyses in epidemiology studies. Cancer
Epidemiology Biomarkers \& Prevention \textbf{25}:407--416.
doi:\href{https://doi.org/10.1158/1055-9965.epi-15-0951}{10.1158/1055-9965.epi-15-0951}.

\leavevmode\hypertarget{ref-Amir2017b}{}%
10. \textbf{Amir A}, \textbf{McDonald D}, \textbf{Navas-Molina JA},
\textbf{Debelius J}, \textbf{Morton JT}, \textbf{Hyde E},
\textbf{Robbins-Pianka A}, \textbf{Knight R}. 2017. Correcting for
microbial blooms in fecal samples during room-temperature shipping.
mSystems \textbf{2}:e00199--16.
doi:\href{https://doi.org/10.1128/msystems.00199-16}{10.1128/msystems.00199-16}.

\leavevmode\hypertarget{ref-Lauber2010}{}%
11. \textbf{Lauber CL}, \textbf{Zhou N}, \textbf{Gordon JI},
\textbf{Knight R}, \textbf{Fierer N}. 2010. Effect of storage conditions
on the assessment of bacterial community structure in soil and
human-associated samples. FEMS Microbiology Letters \textbf{307}:80--86.
doi:\href{https://doi.org/10.1111/j.1574-6968.2010.01965.x}{10.1111/j.1574-6968.2010.01965.x}.

\leavevmode\hypertarget{ref-Song2016}{}%
12. \textbf{Song SJ}, \textbf{Amir A}, \textbf{Metcalf JL},
\textbf{Amato KR}, \textbf{Xu ZZ}, \textbf{Humphrey G}, \textbf{Knight
R}. 2016. Preservation methods differ in fecal microbiome stability,
affecting suitability for field studies. mSystems \textbf{1}:e00021--16.
doi:\href{https://doi.org/10.1128/msystems.00021-16}{10.1128/msystems.00021-16}.

\leavevmode\hypertarget{ref-Costea2017}{}%
13. \textbf{Costea PI}, \textbf{Zeller G}, \textbf{Sunagawa S},
\textbf{Pelletier E}, \textbf{Alberti A}, \textbf{Levenez F},
\textbf{Tramontano M}, \textbf{Driessen M}, \textbf{Hercog R},
\textbf{Jung F-E}, \textbf{Kultima JR}, \textbf{Hayward MR},
\textbf{Coelho LP}, \textbf{Allen-Vercoe E}, \textbf{Bertrand L},
\textbf{Blaut M}, \textbf{Brown JRM}, \textbf{Carton T},
\textbf{Cools-Portier S}, \textbf{Daigneault M}, \textbf{Derrien M},
\textbf{Druesne A}, \textbf{Vos WM de}, \textbf{Finlay BB},
\textbf{Flint HJ}, \textbf{Guarner F}, \textbf{Hattori M},
\textbf{Heilig H}, \textbf{Luna RA}, \textbf{Hylckama Vlieg J van},
\textbf{Junick J}, \textbf{Klymiuk I}, \textbf{Langella P},
\textbf{Chatelier EL}, \textbf{Mai V}, \textbf{Manichanh C},
\textbf{Martin JC}, \textbf{Mery C}, \textbf{Morita H}, \textbf{O'Toole
PW}, \textbf{Orvain C}, \textbf{Patil KR}, \textbf{Penders J},
\textbf{Persson S}, \textbf{Pons N}, \textbf{Popova M}, \textbf{Salonen
A}, \textbf{Saulnier D}, \textbf{Scott KP}, \textbf{Singh B},
\textbf{Slezak K}, \textbf{Veiga P}, \textbf{Versalovic J}, \textbf{Zhao
L}, \textbf{Zoetendal EG}, \textbf{Ehrlich SD}, \textbf{Dore J},
\textbf{Bork P}. 2017. Towards standards for human fecal sample
processing in metagenomic studies. Nature Biotechnology.
doi:\href{https://doi.org/10.1038/nbt.3960}{10.1038/nbt.3960}.

\leavevmode\hypertarget{ref-Haas2011}{}%
14. \textbf{Haas BJ}, \textbf{Gevers D}, \textbf{Earl AM},
\textbf{Feldgarden M}, \textbf{Ward DV}, \textbf{Giannoukos G},
\textbf{Ciulla D}, \textbf{Tabbaa D}, \textbf{Highlander SK},
\textbf{Sodergren E}, \textbf{Methe B}, \textbf{DeSantis TZ},
\textbf{Petrosino JF}, \textbf{Knight R}, \textbf{and BWB}. 2011.
Chimeric 16S rRNA sequence formation and detection in sanger and
454-pyrosequenced PCR amplicons. Genome Research \textbf{21}:494--504.
doi:\href{https://doi.org/10.1101/gr.112730.110}{10.1101/gr.112730.110}.

\leavevmode\hypertarget{ref-Sinha2017}{}%
15. \textbf{Sinha R}, \textbf{Abu-Ali G}, \textbf{Vogtmann E},
\textbf{Fodor AA}, \textbf{Ren B}, \textbf{Amir A}, \textbf{Schwager E},
\textbf{Crabtree J}, \textbf{Ma S}, \textbf{Abnet CC}, \textbf{Knight
R}, \textbf{White O}, \textbf{Huttenhower C}. 2017. Assessment of
variation in microbial community amplicon sequencing by the microbiome
quality control (MBQC) project consortium. Nature Biotechnology.
doi:\href{https://doi.org/10.1038/nbt.3981}{10.1038/nbt.3981}.

\leavevmode\hypertarget{ref-Meisel2016}{}%
16. \textbf{Meisel JS}, \textbf{Hannigan GD}, \textbf{Tyldsley AS},
\textbf{SanMiguel AJ}, \textbf{Hodkinson BP}, \textbf{Zheng Q},
\textbf{Grice EA}. 2016. Skin microbiome surveys are strongly influenced
by experimental design. Journal of Investigative Dermatology
\textbf{136}:947--956.
doi:\href{https://doi.org/10.1016/j.jid.2016.01.016}{10.1016/j.jid.2016.01.016}.

\leavevmode\hypertarget{ref-Caporaso2010}{}%
17. \textbf{Caporaso JG}, \textbf{Lauber CL}, \textbf{Walters WA},
\textbf{Berg-Lyons D}, \textbf{Lozupone CA}, \textbf{Turnbaugh PJ},
\textbf{Fierer N}, \textbf{Knight R}. 2010. Global patterns of 16S rRNA
diversity at a depth of millions of sequences per sample. Proceedings of
the National Academy of Sciences \textbf{108}:4516--4522.
doi:\href{https://doi.org/10.1073/pnas.1000080107}{10.1073/pnas.1000080107}.

\leavevmode\hypertarget{ref-Kozich2013}{}%
18. \textbf{Kozich JJ}, \textbf{Westcott SL}, \textbf{Baxter NT},
\textbf{Highlander SK}, \textbf{Schloss PD}. 2013. Development of a
dual-index sequencing strategy and curation pipeline for analyzing
amplicon sequence data on the MiSeq illumina sequencing platform.
Applied and Environmental Microbiology \textbf{79}:5112--5120.
doi:\href{https://doi.org/10.1128/aem.01043-13}{10.1128/aem.01043-13}.

\leavevmode\hypertarget{ref-Schloss2011}{}%
19. \textbf{Schloss PD}, \textbf{Gevers D}, \textbf{Westcott SL}. 2011.
Reducing the effects of PCR amplification and sequencing artifacts on
16S rRNA-based studies. PLoS ONE \textbf{6}:e27310.
doi:\href{https://doi.org/10.1371/journal.pone.0027310}{10.1371/journal.pone.0027310}.

\leavevmode\hypertarget{ref-Bokulich2012}{}%
20. \textbf{Bokulich NA}, \textbf{Subramanian S}, \textbf{Faith JJ},
\textbf{Gevers D}, \textbf{Gordon JI}, \textbf{Knight R}, \textbf{Mills
DA}, \textbf{Caporaso JG}. 2012. Quality-filtering vastly improves
diversity estimates from illumina amplicon sequencing. Nature Methods
\textbf{10}:57--59.
doi:\href{https://doi.org/10.1038/nmeth.2276}{10.1038/nmeth.2276}.

\leavevmode\hypertarget{ref-Parada2015}{}%
21. \textbf{Parada AE}, \textbf{Needham DM}, \textbf{Fuhrman JA}. 2015.
Every base matters: Assessing small subunit rRNA primers for marine
microbiomes with mock communities, time series and global field samples.
Environmental Microbiology \textbf{18}:1403--1414.
doi:\href{https://doi.org/10.1111/1462-2920.13023}{10.1111/1462-2920.13023}.

\leavevmode\hypertarget{ref-Wang1996}{}%
22. \textbf{Wang GCY}, \textbf{Wang Y}. 1996. The frequency of chimeric
molecules as a consequence of PCR co-amplification of 16S rRNA genes
from different bacterial species. Microbiology \textbf{142}:1107--1114.
doi:\href{https://doi.org/10.1099/13500872-142-5-1107}{10.1099/13500872-142-5-1107}.

\leavevmode\hypertarget{ref-Potapov2017}{}%
23. \textbf{Potapov V}, \textbf{Ong JL}. 2017. Examining sources of
error in PCR by single-molecule sequencing. PLOS ONE
\textbf{12}:e0169774.
doi:\href{https://doi.org/10.1371/journal.pone.0169774}{10.1371/journal.pone.0169774}.

\leavevmode\hypertarget{ref-Kebschull2015}{}%
24. \textbf{Kebschull JM}, \textbf{Zador AM}. 2015. Sources of
PCR-induced distortions in high-throughput sequencing data sets. Nucleic
Acids Research gkv717.
doi:\href{https://doi.org/10.1093/nar/gkv717}{10.1093/nar/gkv717}.

\leavevmode\hypertarget{ref-McInerney2014}{}%
25. \textbf{McInerney P}, \textbf{Adams P}, \textbf{Hadi MZ}. 2014.
Error rate comparison during polymerase chain reaction by DNA
polymerase. Molecular Biology International \textbf{2014}:1--8.
doi:\href{https://doi.org/10.1155/2014/287430}{10.1155/2014/287430}.

\leavevmode\hypertarget{ref-Cline1996}{}%
26. \textbf{Cline J}. 1996. PCR fidelity of pfu DNA polymerase and other
thermostable DNA polymerases. Nucleic Acids Research
\textbf{24}:3546--3551.
doi:\href{https://doi.org/10.1093/nar/24.18.3546}{10.1093/nar/24.18.3546}.

\leavevmode\hypertarget{ref-Acinas2005}{}%
27. \textbf{Acinas SG}, \textbf{Sarma-Rupavtarm R}, \textbf{Klepac-Ceraj
V}, \textbf{Polz MF}. 2005. PCR-induced sequence artifacts and bias:
Insights from comparison of two 16S rRNA clone libraries constructed
from the same sample. Applied and Environmental Microbiology
\textbf{71}:8966--8969.
doi:\href{https://doi.org/10.1128/aem.71.12.8966-8969.2005}{10.1128/aem.71.12.8966-8969.2005}.

\leavevmode\hypertarget{ref-Polz1998}{}%
28. \textbf{Polz MF}, \textbf{Cavanaugh CM}. 1998. Bias in
template-to-product ratios in multitemplate PCR. Applied and
Environmental Microbiology \textbf{64}:3724--3730.

\leavevmode\hypertarget{ref-Brooks2015}{}%
29. \textbf{Brooks JP}, \textbf{David J Edwards}, \textbf{Harwich MD},
\textbf{Rivera MC}, \textbf{Fettweis JM}, \textbf{Serrano MG},
\textbf{Reris RA}, \textbf{Sheth NU}, \textbf{Huang B}, \textbf{Girerd
P}, \textbf{Strauss JF}, \textbf{Jefferson KK}, \textbf{Buck GA}. 2015.
The truth about metagenomics: Quantifying and counteracting bias in 16S
rRNA studies. BMC Microbiology \textbf{15}.
doi:\href{https://doi.org/10.1186/s12866-015-0351-6}{10.1186/s12866-015-0351-6}.

\leavevmode\hypertarget{ref-Suzuki1996}{}%
30. \textbf{Suzuki MT}, \textbf{Giovannoni SJ}. 1996. Bias caused by
template annealing in the amplification of mixtures of 16S rRNA genes by
pcr. Applied and environmental microbiology \textbf{62}:625--630.

\leavevmode\hypertarget{ref-Chandler1997}{}%
31. \textbf{Chandler D}, \textbf{Fredrickson J}, \textbf{Brockman F}.
1997. Effect of pcr template concentration on the composition and
distribution of total community 16S rDNA clone libraries. Molecular
Ecology \textbf{6}:475--482.

\leavevmode\hypertarget{ref-Wagner1994}{}%
32. \textbf{Wagner A}, \textbf{Blackstone N}, \textbf{Cartwright P},
\textbf{Dick M}, \textbf{Misof B}, \textbf{Snow P}, \textbf{Wagner GP},
\textbf{Bartels J}, \textbf{Murtha M}, \textbf{Pendleton J}. 1994.
Surveys of gene families using polymerase chain reaction: PCR selection
and pcr drift. Systematic Biology \textbf{43}:250--261.

\leavevmode\hypertarget{ref-Hansen1998}{}%
33. \textbf{Hansen MC}, \textbf{Tolker-Nielsen T}, \textbf{Givskov M},
\textbf{Molin S}. 1998. Biased 16S rDNA pcr amplification caused by
interference from dna flanking the template region. FEMS Microbiology
Ecology \textbf{26}:141--149.

\leavevmode\hypertarget{ref-Kennedy2014}{}%
34. \textbf{Kennedy K}, \textbf{Hall MW}, \textbf{Lynch MDJ},
\textbf{Moreno-Hagelsieb G}, \textbf{Neufeld JD}. 2014. Evaluating bias
of illumina-based bacterial 16S rRNA gene profiles. Applied and
Environmental Microbiology \textbf{80}:5717--5722.
doi:\href{https://doi.org/10.1128/aem.01451-14}{10.1128/aem.01451-14}.

\leavevmode\hypertarget{ref-Edgar2011}{}%
35. \textbf{Edgar RC}, \textbf{Haas BJ}, \textbf{Clemente JC},
\textbf{Quince C}, \textbf{Knight R}. 2011. UCHIME improves sensitivity
and speed of chimera detection. Bioinformatics \textbf{27}:2194--2200.
doi:\href{https://doi.org/10.1093/bioinformatics/btr381}{10.1093/bioinformatics/btr381}.

\leavevmode\hypertarget{ref-Edgar2017}{}%
36. \textbf{Edgar RC}. 2017. UNBIAS: An attempt to correct abundance
bias in 16S sequencing, with limited success.
doi:\href{https://doi.org/10.1101/124149}{10.1101/124149}.

\leavevmode\hypertarget{ref-Gohl2016}{}%
37. \textbf{Gohl DM}, \textbf{Vangay P}, \textbf{Garbe J},
\textbf{MacLean A}, \textbf{Hauge A}, \textbf{Becker A}, \textbf{Gould
TJ}, \textbf{Clayton JB}, \textbf{Johnson TJ}, \textbf{Hunter R},
\textbf{Knights D}, \textbf{Beckman KB}. 2016. Systematic improvement of
amplicon marker gene methods for increased accuracy in microbiome
studies. Nature Biotechnology \textbf{34}:942--949.
doi:\href{https://doi.org/10.1038/nbt.3601}{10.1038/nbt.3601}.

\leavevmode\hypertarget{ref-Quince2011}{}%
38. \textbf{Quince C}, \textbf{Lanzen A}, \textbf{Davenport RJ},
\textbf{Turnbaugh PJ}. 2011. Removing noise from pyrosequenced
amplicons. BMC Bioinformatics \textbf{12}:38.
doi:\href{https://doi.org/10.1186/1471-2105-12-38}{10.1186/1471-2105-12-38}.

\leavevmode\hypertarget{ref-Salter2014}{}%
39. \textbf{Salter SJ}, \textbf{Cox MJ}, \textbf{Turek EM},
\textbf{Calus ST}, \textbf{Cookson WO}, \textbf{Moffatt MF},
\textbf{Turner P}, \textbf{Parkhill J}, \textbf{Loman NJ},
\textbf{Walker AW}. 2014. Reagent and laboratory contamination can
critically impact sequence-based microbiome analyses. BMC Biology
\textbf{12}.
doi:\href{https://doi.org/10.1186/s12915-014-0087-z}{10.1186/s12915-014-0087-z}.

\leavevmode\hypertarget{ref-Callahan2016}{}%
40. \textbf{Callahan BJ}, \textbf{McMurdie PJ}, \textbf{Rosen MJ},
\textbf{Han AW}, \textbf{Johnson AJA}, \textbf{Holmes SP}. 2016. DADA2:
High-resolution sample inference from illumina amplicon data. Nature
Methods \textbf{13}:581--583.
doi:\href{https://doi.org/10.1038/nmeth.3869}{10.1038/nmeth.3869}.

\leavevmode\hypertarget{ref-Edgar2016}{}%
41. \textbf{Edgar RC}. 2016. UNOISE2: Improved error-correction for
illumina 16S and ITS amplicon sequencing.
doi:\href{https://doi.org/10.1101/081257}{10.1101/081257}.

\leavevmode\hypertarget{ref-Amir2017a}{}%
42. \textbf{Amir A}, \textbf{McDonald D}, \textbf{Navas-Molina JA},
\textbf{Kopylova E}, \textbf{Morton JT}, \textbf{Xu ZZ},
\textbf{Kightley EP}, \textbf{Thompson LR}, \textbf{Hyde ER},
\textbf{Gonzalez A}, \textbf{Knight R}. 2017. Deblur rapidly resolves
single-nucleotide community sequence patterns. mSystems \textbf{2}.
doi:\href{https://doi.org/10.1128/msystems.00191-16}{10.1128/msystems.00191-16}.

\leavevmode\hypertarget{ref-Schloss2018}{}%
43. \textbf{Schloss PD}. 2018. Identifying and overcoming threats to
reproducibility, replicability, robustness, and generalizability in
microbiome research. mBio \textbf{9}.
doi:\href{https://doi.org/10.1128/mbio.00525-18}{10.1128/mbio.00525-18}.

\leavevmode\hypertarget{ref-Seekatz2016}{}%
44. \textbf{Seekatz AM}, \textbf{Rao K}, \textbf{Santhosh K},
\textbf{Young VB}. 2016. Dynamics of the fecal microbiome in patients
with recurrent and nonrecurrent clostridium difficile infection. Genome
Medicine \textbf{8}.
doi:\href{https://doi.org/10.1186/s13073-016-0298-8}{10.1186/s13073-016-0298-8}.

\leavevmode\hypertarget{ref-Schloss2009}{}%
45. \textbf{Schloss PD}, \textbf{Westcott SL}, \textbf{Ryabin T},
\textbf{Hall JR}, \textbf{Hartmann M}, \textbf{Hollister EB},
\textbf{Lesniewski RA}, \textbf{Oakley BB}, \textbf{Parks DH},
\textbf{Robinson CJ}, \textbf{Sahl JW}, \textbf{Stres B},
\textbf{Thallinger GG}, \textbf{Horn DJV}, \textbf{Weber CF}. 2009.
Introducing mothur: Open-source, platform-independent,
community-supported software for describing and comparing microbial
communities. Applied and Environmental Microbiology
\textbf{75}:7537--7541.
doi:\href{https://doi.org/10.1128/aem.01541-09}{10.1128/aem.01541-09}.

\leavevmode\hypertarget{ref-Rognes2016}{}%
46. \textbf{Rognes T}, \textbf{Flouri T}, \textbf{Nichols B},
\textbf{Quince C}, \textbf{Mahé F}. 2016. VSEARCH: A versatile open
source tool for metagenomics. PeerJ \textbf{4}:e2584.
doi:\href{https://doi.org/10.7717/peerj.2584}{10.7717/peerj.2584}.

\leavevmode\hypertarget{ref-Westcott2017}{}%
47. \textbf{Westcott SL}, \textbf{Schloss PD}. 2017. OptiClust, an
improved method for assigning amplicon-based sequence data to
operational taxonomic units. mSphere \textbf{2}:e00073--17.
doi:\href{https://doi.org/10.1128/mspheredirect.00073-17}{10.1128/mspheredirect.00073-17}.

\leavevmode\hypertarget{ref-r_citation_2018}{}%
48. \textbf{R Core Team}. 2018. R: A language and environment for
statistical computing. R Foundation for Statistical Computing, Vienna,
Austria.

\leavevmode\hypertarget{ref-Tidyverse2017}{}%
49. \textbf{Wickham H}. 2017. Tidyverse: Easily install and load the
'tidyverse'.

\leavevmode\hypertarget{ref-Vegan2018}{}%
50. \textbf{Oksanen J}, \textbf{Blanchet FG}, \textbf{Friendly M},
\textbf{Kindt R}, \textbf{Legendre P}, \textbf{McGlinn D},
\textbf{Minchin PR}, \textbf{O'Hara RB}, \textbf{Simpson GL},
\textbf{Solymos P}, \textbf{Stevens MHH}, \textbf{Szoecs E},
\textbf{Wagner H}. 2018. Vegan: Community ecology package.

\newpage

\textbf{Figure 1. The error rate of assembled sequence reads increases
with the number of rounds of PCR used and follows the relative error
rates provided by the manufacturers; however, much of this error is
mediated by denoising.}

\includegraphics[width=0.9\columnwidth]{../results/figures/mock_error.pdf}

\newpage

\textbf{Figure 2. The fraction of all denoised sequences that were
identified as being chimeric increases with the number of rounds of PCR
used and varied between polymerases.} (A) Sequencing of a mock community
allowed us to identify the total fraction of sequences that were
chimeric as well as the sensitivity and specificity of UCHIME to detect
those chimeras. (B) Sequencing of four human stool samples after using
one of five different polymerases again demonstrated increased rate of
chimera formation with increasing number of rounds of PCR and variation
across polymerases.

\includegraphics[width=0.9\columnwidth]{../results/figures/chimera_plots.pdf}

\newpage

\textbf{Figure 3. The relative abundances of reads mapped to reference
sequences differed subtly from the expected relative abundances as
determined by shotgun metagenomic sequencing.} Bias did not increase
with number of rounds of PCR or vary by polymerase or the guanine and
cytosine content of the fragment. The expected relative abundance of
each organism is indicated by the horizontal gray line. The percentage
of bases that were guanines or cytosines within the V4 region of the 16S
rRNA genes in each organism is indicated by the number in the lower left
corner of each panel.

\includegraphics[width=0.9\columnwidth]{../results/figures/species_bias.pdf}

\newpage

\textbf{Figure 4. Despite evidence of subtle PCR bias at the genome
level, there was significant evidence of bias using community-wide
metrics that grew with the number of rounds of PCR when using a mock
community.} (A) With the exception of data collected using the Kappa
polymerase, the OTU richness and Shannon diversity values increased with
number of rounds of PCR and the inclusion of residual sequencing errors
and chimeras. The horizontal black line indicates the expected richness
and diversity if there were no bias, errors, or chimeras. (B) Relative
to the mock community sampled after 25 rounds of PCR, the Bray-Curtis
distance to the communities sampled after 30 and 35 rounds of PCR
increased for all polymerases. (C) The variation between samples
collected after 20, 25, 30, and 35 rounds of PCR for the five
polymerases demonstrated a significant change in the community driven by
the number of rounds of PCR and the polymerase used.

\includegraphics[width=0.8\columnwidth]{../results/figures/mock_community.pdf}

\newpage

\textbf{Figure 5. Sequencing of human stool samples indicated clear
increase in bias with number of rounds of PCR, however, the bias
appeared to be consistent within each sample.} (A) With the exception of
data collected using the Kappa polymerase, the OTU richness and Shannon
diversity values increased with number of rounds of PCR. (B) Relative to
the stool communities sampled after 25 rounds of PCR, the Bray-Curtis
distance to the stool communities sampled after 30 and 35 rounds of PCR
was inconsistent and there was little difference in variation for data
collected using the Kappa polymerase. (C) The variation between stool
samples was larger than the amount of variation introduced by varying
the number of rounds of PCR or polymerase.

\includegraphics[width=0.9\columnwidth]{../results/figures/stool_community.pdf}

\newpage

\textbf{Figure 6. There is little evidence for PCR drift adversely
impacting the results of amplicon studies.} The average distance between
replicates of sequencing the same mock community or between the the
human stool samples did not vary by number of rounds of PCR or by
polymerase.

\includegraphics[width=0.9\columnwidth]{../results/figures/drift.pdf}

\newpage

\textbf{Figure S1: With the exception of the seuqence data generated
using the Kappa polymerase, the ratio of the two \emph{Salmonella
enterica} V4 sequences to a value was lower than the expected ratio of
6:1.} The two \emph{S. enterica} V4 sequences differed by a single base.

\includegraphics[width=0.9\columnwidth]{../results/figures/salmonella_bias.pdf}

\end{document}